### 2020-10-13 README.Rmd
Final edits

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r Title, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Family Fortunes: A Quantitative Analysis of the Early Iron Age Cemeteries of Knossos"
```

# knossoscemeteries

This repository contains the data and code for the paper:

> Pollard, Dominic (2020). _`r Title`_. Name of journal/book, <https://doi.org/xxx/xxx>

The pre-print available is online here:

> Pollard, Dominic (2020). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Repository contents

All data, text and code to recreate the paper and analysis therein are found in the 'analysis' folder. Within this, 'analysis/data' contains all raw and derived data, in .csv format. All charts and images from the paper can be found in 'analysis/figures', while the paper itself, and the BibTex file containing all cited works are found in 'analysis/paper'. The folder 'analysis/templates' contains the citation style file, and a template file used to set the style for the Word document produced by 'Family_Fortunes_Article.Rmd'.

### How to cite

Please cite this compendium as:

> Pollard, Dominic (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/DCPollard94/knossoscemeteries/archive/master.zip>

Or you can install this compendium as an R package from GitHub with:
```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("DCPollard94/knossoscemeteries")
```
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.